---
title: "Bio4158 - Devoir 4"
author: "Groupe XX - Nom1, Nom2, Nom3"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    number_sections: false
---


# The problem

The peppered moth (_Biston betularia_ Lepidoptera : Geometridae) is a nocturnal insect common in Europe and North America. Adults fly at night and stay put during the day on birch trees. Adult coloration is variable, ranging from a clear form (typica) to a melanic form that is very dark (carbonaria), with many intermediate forms.

The proportion of individuals of each form has varied considerably in natural populations over the last 150 years. The melanic form was rare before 1850 but represented over 90% of the population in industrial sectors of England by 1900. According to Majerus (1998), J.W. Tutt was the first to suggest in 1896 an hypothesis to explain the temporal change in the proportion of the dark form. 

__Biological hypothesis__: Differences in the relative frequency of the two forms arise because survival from bird predation of each depends on habitat color. In particular, the pale form is less visible in non-polluted environments when birch trees are covered by pale greyish lichens. Inversely, the dark form is favored in industrial areas where soot darkens their resting places. Therefore, the increase of the dark form (carbonaria) in England resulted from bird predation and a change in habitat (soot pollution). 

# The study

In a study designed to test this hypothesis, a random sample of 30 insectivorous birds was used to estimate predation susceptibility of the pale and dark forms on both pale and dark substrates. Two growth chamber were set up containing branches and leaves:

1. in the first, the branches and leaves were taken from a polluted area and were covered with soot (and were thus dark);
2. in the second, the branches and leaves were taken from a non-polluted area and were thus lighter in colour.

Each bird was introduced in turn into each chamber (half into the dark first and half into the light first), with a fasting period in between to keep hunger constant. Just before introducing a bird to a chamber, 50 pale moths (typica) and 50 dark moths (carbonaria) were randomly selected and put in the chamber. After a fixed period of time, the mean number of pale and dark moths captured by the bird in the pale and dark chambers was determined. After each experiment, the remaining moths were discarded before putting in fresh moths for the next experiment.


# The data

The file `Biston_student.csv` contains a fraction of the data that a professor has managed to unearth after his MSc student left without a trace prior to publishing their results. Unfortunately, the data sheets are incomplete and do not allow him to tell which value comes from which bird. In this file there are 30 lines (representing the 30 birds) and two columns (one for the number of dark moths captured in the dark chamber and one for the number of dark moths captured in the pale chamber). However, numbers on a given line do NOT necessarily come from the same bird (i.e. the order of observations in the two columns has been mixed up).

The file `Biston_pd2.csv` is later reconstructed by a post-doc from the same lab. This post-doc is much more persistent than the prof and manages to find additional data sheets that allow her to match each result to a particular bird. This file has the same structure as `Biston_student.csv`, and contains the same data, except that observations on the same line now come from the same bird (i.e. paired observations are known). 
```{r}
# R code

library("car")
library("lmtest")
library("boot")
library("pwr")
library("ggplot2")
library("performance")
library(tidyverse)
library(lsr)
```


# The task

1. Analyze the data and determine if the results in `Biston_student.csv` support or refute the biological hypothesis. 
   a. a statement of the biological hypothesis and its prediction (1 point)
   -the dark moths will have a higher survival rate in the dark room than in the pale room
    
   b. a statement of the null hypothesis (1 point)
   -the average amount of dark moths eaten in the dark room will be equal to the average amount of dark moths eaten in the pale room
   c. fit the adequate statistical model, and evaluate the statistical assumptions
```{r}
b_student <- read.csv("C://Users//kinng//Desktop//university//r code location year 4//bio 4158 lab 1//Biston_student.csv")
pd_2<-read.csv("C://Users//kinng//Desktop//university//r code location year 4//bio 4158 lab 1//Biston_pd_2.csv")

mygraph <- ggplot(
  data = b_student,
  aes(x = number)
) +
  xlab("number")
# add data to the mygraph ggplot
mygraph <- mygraph +
  geom_density() + # add data density smooth
  geom_rug() + # add rug (bars at the bottom of the plot)
  geom_histogram( # add black semitransparent histogram
    aes(y = ..density..),
    color = "purple", alpha = 0.3
  ) +
  # add normal curve in red, with mean and sd from fklength
  stat_function(
    fun = dnorm,
    args = list(
      mean = mean(b_student$number),
      sd = sd(b_student$number)
    ),
    color = "red"
  )
mygraph + facet_grid(. ~ room)

```
```{r}
modelgraph <- lm(number ~ room, data = b_student)
par(mfrow = c(2, 2))
plot(modelgraph)


```
```{r}
leveneTest(modelgraph)
```

Here, we calculate the mean and the standard deviation for the dark room and the pale room 

```{r}

b_student %>% group_by (room)%>%
          summarise(dark_moth_mean = mean(number), dark_moth_sd=sd(number))
```
 
As we can see above, the standard deviations are relatively similar, and when checking the levene test, we see that it fails to reject the null where the variances are equal.


based on the QQ plot, we can see that the data follows a relatively normal distribution(the population sample is large enough that the student t test should still be strong). Furthermore, the spread of the residuals is even in between the two rooms, giving a visual confirmation of homoskedasticity.
```{r}
t.test(
  number ~ room,
  data = b_student,
  alternative = "two.sided",
  var.equal = TRUE
)
```



   d. The statistical conclusion (i.e. inference). Do you accept or reject the null hypothesis? Include your test results. (3 points)
   
   **based on the above t test, the p value(=0.1258) is greater than our alpha of 0.05, as a result we fail to reject the null hypothesis that the average amount of moths eaten in the dark and pale rooms are the statistically different.**

  e. The “biological” conclusion (i.e. a qualified statement about the biological hypothesis concerning moth survival, including an estimate of raw effect size, its precision, and whether an effect of this magnitude seems biologically interesting) (2 points). With respect to the latter, keep in mind that even small differences in survival can compound across generations to result in substantial evolutionary change given enough time (and assuming that colour variation is heritable).
  
  

  
  
```{r}
dark_moth_mean = 29.4667
pale_moth_mean = 32.2667
raw_effect_size = pale_moth_mean-dark_moth_mean
raw_effect_size

dark_s = 7.074643	
pale_s = 6.887937


```
we have the raw effective size to be *2.8*. And we can caclulate the Cohden effect size with this code below

```{r}

group1 <- (b_student %>% filter(room=="dark"))$number
group2 <- (b_student %>% filter(room=="pale"))$number

#calculate Cohen's d
cohensD(group1, group2)


```

The adjusted effect size is **d= 0.401**, which will be used in our calculation for power 

Based on this efect size, we can see that there is at least 3more dark moths eaten in the pale rooms than in the dark rooms. This is a **biologically significant difference** with a confidence interval of **(-6.4085412,0.8085412)**. Although 0 falls within the interval, we can see that the CI skews towards having more moths in dark rooms surviving than in pale rooms. Compared to the mean, the confidence interval is quite large, which implies a low level of precision. 

Now we need to assees the power of the model 


```{r}

pwr.t.test(
  n=30,
  d=0.401,
  sig.level=0.05,
  type= "two.sample"
  
)
  

```
  
we can see that the power is **0.333** which is very low.
  
  f. A statement about your confidence in the statistical conclusion (for example with respect to the violation of certain assumptions, to potential biases created by missing data, to suboptimal experimental design, and/or to power issues ) (2 points)
  
 **come write later** 
  
  
  
  
  
  
2. Analyze the data and determine if the results in `Biston_pd2.csv `support or refute the biological hypothesis.
    a. a statement of the biological hypothesis and its prediction (1 point)
    -the dark moths will have a higher survival rate in the dark room than in the pale room
    
  b. a statement of the null hypothesis (1 point)
    -the average amount of moths that survive in the pale room is equal to the amount of moths that survive in the dark room
  c. fit the adequate statistical model, and evaluate the statistical assumptions   
```{r}
biston_pd2<-pd_2

biston_pd2$dark_pale<-c(biston_pd2$dark - biston_pd2$pale)

## divide the display 
par(mfrow=c(1,2))

mygraph2 <- ggplot(
  data = biston_pd2,
  aes(x = pale)
) +
  xlab("Dark-Pale")
# add data to the mygraph ggplot
mygraph2 <- mygraph2 +
  geom_density() + # add data density smooth
  geom_rug() + # add rug (bars at the bottom of the plot)
  geom_histogram( # add black semitransparent histogram
    aes(y = ..density..),
    color = "green", alpha = 0.3
  ) +
  # add normal curve in red, with mean and sd from fklength
  stat_function(
    fun = dnorm,
    args = list(
      mean = mean(biston_pd2$dark_pale),
      sd = sd(biston_pd2$dark_pale)
    ),
    color = "red"
  )







############################
  mygraph3 <- ggplot(
  data = biston_pd2,
  aes(x = dark)
) +
  xlab("Dark")
# add data to the mygraph ggplot
mygraph3 <- mygraph3 +
  geom_density() + # add data density smooth
  geom_rug() + # add rug (bars at the bottom of the plot)
  geom_histogram( # add black semitransparent histogram
    aes(y = ..density..),
    color = "green", alpha = 0.3
  ) +
  # add normal curve in red, with mean and sd from fklength
  stat_function(
    fun = dnorm,
    args = list(
      mean = mean(biston_pd2$dark),
      sd = sd(biston_pd2$dark)
    ),
    color = "red"
  )


mygraph2



```

from the graphs, we can visually assume that the data is normally distributed, however we still need to check the residual assumptions.


```{r}

modelgraph2 <- lm(dark_pale~bird, data = biston_pd2)
par(mfrow = c(2, 2))
plot(modelgraph2)

```
```{r}

shapiro.test(residuals(modelgraph2))


```

based on the pvalue of the shapiro-wilk test we can conclude that the deviation from normality of the difference between the dark and light rooms are not significant.

```{r}

t.test(biston_pd2$dark,biston_pd2$pale, 
       alternative = "two.sided",
       paired = TRUE)
```

  
  d. The statistical conclusion (i.e. inference). Do you accept or reject the null hypothesis? Include your test results. (3 points)
  
  since the pvalue(=0.045) is less than the alpha=0.05, we can reject the null and conclude that there is a statistically significant difference bet ween the pale and dark rooms.
  
  e. The “biological” conclusion (i.e. a qualified statement about the biological hypothesis concerning moth survival, including an estimate of raw effect size, its precision, and whether an effect of this magnitude seems biologically interesting) (2 points). With respect to the latter, keep in mind that even small differences in survival can compound across generations to result in substantial evolutionary change given enough time (and assuming that colour variation is heritable).
  
  
  
  f. A statement about your confidence in the statistical conclusion (for example with respect to the violation of certain assumptions, to potential biases created by missing data, to suboptimal experimental design, and/or to power issues ) (2 points)
3. Assuming that a difference of 5% (or 2.5 moths) of predation rate between treatments is biologically meaningful, what would be the power to detect this with alpha=0.05 and the data collected by:
    a. the student?
    b. the postdoc?
    c. Comment the difference between the power using the student and postdoc data.

